Specialized Cost Function for Maximizing Body Area Network Lifetime Using Global Routing Algorithms\\ Masters Thesis

### 2011-09-16 docs/thesis/thesis.tex
Re-formatted thesis.

\documentclass{article}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\begin{document}
\title{Specialized Cost Function for Maximizing Body Area Network Lifetime Using Global Routing Algorithms\\
Masters Thesis}
\author{Alvaro Prieto\\
BS/MS Electrical Engineering\\
\texttt{axp8664@rit.edu}}
\maketitle

@@

This work primarily focuses on maximizing the lifetime of WBANs. It also includes a basic software framework for developing WBANs, a sample Wireless Electrocardiogram (ECG) application, and a simple link cost algorithm development platform.

\section{Summary of Contributions}
The following is a list of contributions presented in this work.
\begin{itemize}
\item \textbf{Software platform for WBAN research.} - The software platform was developed using only free and open source software (FOSS) tools. Radio libraries for the Texas Instruments (TI) CC430 device were ported over from TI reference code, while libraries for the CC2500 radio were written from scratch. Other features include timer control for scheduling and time-stamping, serial communication, some digital I/O, and analog input. Included in the platform were limited software tools for data capture and visualization on a host computer.

@@

@@

Figures~\ref{fig:walk2-c0}~and~\ref{fig:walk2-c100}, on the other hand, show an increase in network lifetime. For example, at the end of the 5 minute sampling run with $M=0$, the device which consumed the most power was ED7 with approximately 1.7 mW. On the other hand, with $M=100$, all EDs consume power in a uniform fashion and end up with less than 1.3 mW each. One important observation is that the network with $M=0$ consumed less overall power (~9.6mW) than the one with $M=100$ (~10.1mW). This is expected, due to the fact that the new cost function specifically avoids the most efficient path for a specific node in favor of the most efficient path for the whole network. While more energy is consumed overall, no single ED consumes significantly more power.

@@

\section{Conclusion}
The developed WBAN software libraries were successfully used to implement both the wireless ECG system and the full-body routing system. The wireless ECG system was able to sample data at a rate of 300Hz and display it in the host computer. The routing algorithm testing platform facilitated the development of the new link cost function to maximize network lifetime.

After analyzing the results, the new link cost function successfully improves network lifetime. The new cost function outperformed the traditional one in trials dealing with a dynamic environment, where the subject was walking. When the subject was not moving, both link costs performed about the same. Note that, in this case, performance is defined with respect to network lifetime, and not overall power consumption.

@@

%\listoftables

>>>>

\documentclass{article}

\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{fullpage}
\usepackage[top=1.25in, bottom=1.25in, left=1.5in, right=1.0in]{geometry}

\usepackage{setspace}
\doublespacing

%\addtolength{\oddsidemargin}{0.5in}
%\addtolength{\evensidemargin}{0.5in}
%\addtolength{\topmargin}{0.25in}
%\addtolength{\textwidth}{-0.5in}

\begin{document}
\title{Specialized Cost Function for Maximizing Body Area Network Lifetime Using Global Routing Algorithms\\
Masters Thesis}
\author{Alvaro Prieto\\
BS/MS Electrical Engineering\\
Electrical and Microelectronic Engineering Department\\
Rochester Institute of Technology\\
Rochester, NY\\
\texttt{axp8664@rit.edu}}

\thispagestyle{empty}
\maketitle

@@

Presented in this work is a novel global routing algorithm link-cost function to maximize network lifetime in WBANs. Also included are a basic software framework for developing WBANs, a sample Wireless Electrocardiogram (ECG) application, and a simple link cost algorithm development platform.

\section*{Summary of Contributions}
The following is a list of contributions presented in this work.
\begin{itemize}
\item \textbf{Software platform for WBAN research.} - The software platform was developed using only free and open source software (FOSS) tools. Radio libraries for the Texas Instruments (TI) CC430 device were ported over from TI reference code, while libraries for the CC2500 radio were written from scratch. Other features include timer control for scheduling and time-stamping, serial communication, some digital I/O, and analog input. Included in the platform were limited software tools for data capture and visualization on a host computer.

@@

\item \textbf{Publication} A. Prieto, G. R. Tsouri, "Efficient Global Routing Algorithm for Prolonging Network Lifetime in Body Area Networks", in preparation IEEE Wireless Communication Letters.

@@

Figures~\ref{fig:walk2-c0}~and~\ref{fig:walk2-c100}, on the other hand, show an increase in network lifetime. For example, at the end of the 5 minute sampling run with $M=0$, the device which consumed the most power was ED7 with approximately 1.7 mW. On the other hand, with $M=100$, all EDs consume power in a uniform fashion and end up with less than 1.3 mW each. One important observation is that the network with $M=0$ consumed less overall power (~9.6mW) than the one with $M=100$ (~10.1mW). This is expected, due to the fact that the new cost function specifically avoids the most efficient path for a specific node in favor of the most efficient path for the whole network. While more energy is consumed overall, no single ED consumes much more power than others.

Another way of interpreting these results is that energy use in the network is more efficient. When a single ED depletes its energy source using the traditional link cost, there is a significant amount of energy left unused in the network. At the same time, when an ED dies in a network using the new link cost, there should be much less energy left in the rest of the network.

@@

\section{Conclusion}
The developed WBAN software libraries were successfully used to implement both the wireless ECG system and the full-body routing system. The wireless ECG system was able to sample data at a rate of 300Hz and display in real-time through the host computer. The routing algorithm testing platform facilitated the development of the new link cost function to maximize network lifetime.

After analyzing the results, the new link cost function successfully improved network lifetime. The new cost function outperformed the traditional one in trials dealing with a dynamic environment, where the subject was walking. When the subject was not moving, both link costs performed about the same. Note that, in this case, performance is defined with respect to network lifetime, and not overall power consumption.

@@
